---
title: "Vincent's part on SAS Project"
author: "Vincent William Hadiasali"
date: "`r Sys.Date()`"
output: html_document
---

Libraries
```{r}
library(ggplot2)
library(tseries)
library(forecast)
library(dplyr)
library(lubridate)
library(xts)
library(zoo)
```

Import necessary files

```{r}
order <- read.csv("~/University File/Year 3 Spring/MA409/Project/OlistData/olist_order_items_dataset.csv")
status <- read.csv("~/University File/Year 3 Spring/MA409/Project/OlistData/olist_orders_dataset.csv")
payment <- read.csv("~/University File/Year 3 Spring/MA409/Project/OlistData/olist_order_payments_dataset.csv")
product <- read.csv("~/University File/Year 3 Spring/MA409/Project/OlistData/olist_products_dataset.csv")
translation <- read.csv("~/University File/Year 3 Spring/MA409/Project/OlistData/product_category_name_translation.csv")
```

If the order is canceled or unavailable, that means it cannot be counted to revenue, sales number and even freight value.
```{r}
# Match the product categorical name by the product ID
verified <- status[status$order_status != "canceled"
                   & status$order_status != "unavailable",]
verified_order <- merge(order, verified, by = "order_id")
```

Combined the data and use the translated product category name.
Since the analysis says to 2018, erase anything after 2018.
```{r}
combined_data <- merge(verified_order, product, by = "product_id")
translated_data <- merge(combined_data, translation, by = "product_category_name")
# Sort based on the date
df <- translated_data[order(translated_data$shipping_limit_date),
                    c("order_id", "product_category_name_english", "shipping_limit_date", "order_item_id", "freight_value")]
# Only keep the necessary details of time
df$shipping_limit_date <- as.Date(df$shipping_limit_date)
df <- df[df$shipping_limit_date < as.Date("2019-01-01"),]
nrow(df)
```

Monthly revenue trend
```{r}
revenue_df <- merge(df, payment, by = "order_id")
revenue <- aggregate(revenue_df$payment_value, list(revenue_df$shipping_limit_date), FUN=sum)
revenue$Month <- floor_date(revenue$Group.1, "month")
m_revenue <- revenue %>% group_by(Month) %>% summarize(sum = sum(x))
m_revenue <- m_revenue[-c(2, 23), ]
plot(m_revenue, type = 'l', main = "Trend of total revenue by month", ylab = "Total revenue")
```

Monthly total number of order
```{r}
order_df <- aggregate(df$order_item_id, list(df$shipping_limit_date), FUN=sum)
order_df$Month <- floor_date(order_df$Group.1, "month")
m_order <- order_df %>% group_by(Month) %>% summarize(sum = sum(x))
m_order <- m_order[-c(1, 3, 24), ]
plot(m_order, type = 'l', main = "Trend of total order number by month",
     ylab = "Total number of order")
```

freight value prediction analysis

Before freight value time series prediction, we must convert it to monthly
```{r}
freight <- aggregate(df$freight_value, list(df$shipping_limit_date), FUN=sum)

#round dates down to month
freight$month <- floor_date(freight$Group.1, "month")

#find mean sales by month
m_freight <- freight %>% group_by(month) %>% summarize(sum = sum(x))
m_freight

plot(m_freight, type = 'o')
```

Because there are three months where most of days are missing, we cannot include them in forecasting.
```{r}
m_freight <- m_freight[-c(1, 3, 24), ]
m_freight

adf.test(m_freight$sum)
Box.test(m_freight$sum, type="Ljung-Box")
plot(forecast(m_freight$sum, h = 120))
auto.arima(m_freight$sum)

#Differencing
adf.test(diff(m_freight$sum))
Box.test(diff(m_freight$sum), type="Ljung-Box")
plot(forecast(diff(m_freight$sum), h = 120), ylim = c(-30000, 50000))

#Log transformation
adf.test(log(m_freight$sum))
Box.test(log(m_freight$sum), type="Ljung-Box")
plot(forecast(log(m_freight$sum), h = 120))

#Differencing and Log transformation
adf.test(diff(log(m_freight$sum)))
Box.test(diff(log(m_freight$sum), type="Ljung-Box"))
plot(forecast(diff(log(m_freight$sum)), h = 120))
```

