---
title: "FTS Project"
author: "Vincent William Hadiasali"
date: "`r Sys.Date()`"
output: html_document
---
# Overview
```{r library}
library(readr)
library(tseries)
library(fGarch)
library(rugarch)
library(stats)
library("FinTS")
library(forecast)
library(tsDyn)
library(urca)
```


```{r}
data1 <- read.csv("C:/Users/vince/Documents/University File/Year 4 Fall/FIN5017/Project/data1.csv")
data1$Date <- as.Date(data1$Date, format = "%m/%d/%y")
data2 <- read.csv("C:/Users/vince/Documents/University File/Year 4 Fall/FIN5017/Project/data2.csv")
data2$Date <- as.Date(data2[,2], format = "%m/%d/%y")
```

# Time Series Analysis on the future prices of 4 foods.

## Basic Time Series plotting for all future price

```{r pressure, echo=FALSE}
ncoldata1 <- ncol(data1)
colnames(data1) <- gsub("[_.]", " ", colnames(data1))


# Loop to create plots for each column in the data frame (except the first column)
for (i in 2:ncol(data1)) { # Start from the 2nd column, assuming the 1st is 'Date'
  plot(
    data1[, i],                       # The i-th column of the data frame
    ylab = "Value",                   # Y-axis label
    xlab = "Index",                   # X-axis label (adjust if 'Date' is meaningful)
    main = colnames(data1)[i],        # Use the column name as the title
    type = "l"                        # Line with points
  )
}
```
## Stationarity test
```{r}
#Test its stationarity
for (i in 2:ncol(data1)) { # Start from the 2nd column, assuming the 1st is 'Date'
  cat("The ADF Test result of", colnames(data1)[i])
  adf_result <- adf.test(data1[, i], alternative = "stationary")
  
  # Print the test result
  print(adf_result)                       # Line with points
}
```


## GARCH Test
```{r}

# 进行LM检验
# Loop through columns (excluding the first column, assuming it's 'Date')
for (i in 2:ncol(data1)) {
  # Fit a simple linear model (or replace this with your preferred model)
  t <- 1:nrow(data1)
  model <- lm(data1[, i] ~ t)  # This fits a model with just an intercept
  
  # Extract residuals from the model
  res <- residuals(model)
  
  # Perform ARCH test on the residuals
  arch_test_result <- ArchTest(res, , FALSE)
  
  # Print results
  cat("ARCH Test for column:", colnames(data1)[i], "\n")
  print(arch_test_result)
  cat("\n")
}

```




## BDS Test for Non-linear model detection
```{r}
for (i in 2:ncol(data1)) {
    cat("\nBDS Test for column:", colnames(data1)[i], "\n")  # Print column name
    tryCatch(
      {
        # Print the test result
        print(bds.test(data1[, i]))
      },
      error = function(e) {
        # Handle errors (e.g., invalid input for bds.test)
        cat("Error in BDS Test for column:", colnames(data1)[i], "\n")
        print(e$message)
      }
    )
  }
```

## Test SETAR Model on each future price
```{r}
library(tseries)  # Load the required library for bds.test()

# Loop through columns (excluding the first, assuming it's an index or 'Date')
for (m in 2:5) {
  cat("for m = ", m, "\n")
  for (i in 2:ncol(data1)) {
    cat("\nSETAR Test for column:", colnames(data1)[i], "\n")  # Print column name
    tryCatch(
      {
        # Print the test result
        print(setarTest(data1[, i], m = m))
      },
      error = function(e) {
        # Handle errors (e.g., invalid input for bds.test)
        cat("Error in BDS Test for column:", colnames(data1)[i], "\n")
        print(e$message)
      }
    )
  }
}
```

## Fit the SETAR Model based on the SETAR Test, get its threshold and plot them
### Soya Bean

```{r}
library(ggplot2)

setar_bean <- setar(data1[, 2], m = 2, thDelay = 1)
summary(setar_bean)

regime_bean <- ifelse(data1[, 2] <= 3831, "Low Regime", "High Regime")

ggplot(data1, aes(x = 1:nrow(data1), y = data1[,2])) +
  geom_line() +
  geom_hline(yintercept = 3831, color = "red", linetype = "dashed") +
  geom_point(aes(color = regime_bean)) +
  labs(title = "Soya Bean Meal Future Price: Regimes", y = "Price", x = "Index") +
  scale_color_manual(values = c("Low Regime" = "blue", "High Regime" = "green"))
```

### Rice
```{r}
library(ggplot2)

setar_rice <- setar(data1[, 3], m = 5, thDelay = 1)
summary(setar_rice)

regime_rice <- ifelse(data1[, 3] <= 3440, "Low Regime", "High Regime")

ggplot(data1, aes(x = 1:nrow(data1), y = data1[,3])) +
  geom_line() +
  geom_hline(yintercept = 3440, color = "red", linetype = "dashed") +
  geom_point(aes(color = regime_rice)) +
  labs(title = "Rice Future Price: Regimes", y = "Price", x = "Index") +
  scale_color_manual(values = c("Low Regime" = "blue", "High Regime" = "green"))
```

### Wheat
```{r}
setar_wheat <- setar(data1[, 5], m = 3, thDelay = 2) #pick such thDelay because of lower AIC
summary(setar_wheat)

regime_wheat <- ifelse(data1[, 5] <= 852.5, "Low Regime", "High Regime")



ggplot(data1, aes(x = 1:nrow(data1), y = data1[,5])) +
  geom_line() +
  geom_hline(yintercept = 852.5, color = "red", linetype = "dashed") +
  geom_point(aes(color = regime_wheat)) +
  labs(title = "Wheat Future Price: Regimes", y = "Price", x = "Index") +
  scale_color_manual(values = c("Low Regime" = "blue", "High Regime" = "green"))
```
# Log Return
## Transformation
```{r}
# Assuming data1 is your dataset, and columns 2 to 5 contain the price data
log_r_future <- data1  # Create a new data frame to store log-returns

# Loop through each column (2 to 5)
for (col in 2:5) {
  log_r_future[, col] <- c(NA, diff(log(data1[, col])))
}
colnames(log_r_future)[2:5] <- paste("Log return for", colnames(log_r_future)[2:5])
  
# View the computed log-returns
log_r_future <- log_r_future[-1, ]
head(log_r_future)
```

```{r}
for (i in 2:ncol(log_r_future)) { # Start from the 2nd column, assuming the 1st is 'Date'
  plot(
    log_r_future[,i],                       # The i-th column of the data frame
    ylab = "Value",                   # Y-axis label
    xlab = "Index",                   # X-axis label (adjust if 'Date' is meaningful)
    main = colnames(log_r_future)[i],        # Use the column name as the title
    type = "l"                        # Line with points
  )
}
```

```{r}
for (i in 2:ncol(data1)) { # Start from the 2nd column, assuming the 1st is 'Date'
  cat(colnames(log_r_future)[i])
  adf_result <- adf.test(log_r_future[, i], alternative = "stationary")
  
  # Print the test result
  print(adf_result)                       # Line with points
}
```

## GARCH Test
```{r}

# 进行LM检验
# Loop through columns (excluding the first column, assuming it's 'Date')
for (i in 2:ncol(log_r_future)) {
  # Fit a simple linear model (or replace this with your preferred model)
  t <- 1:nrow(log_r_future)
  log_model <- lm(log_r_future[, i] ~ t)  # This fits a model with just an intercept
  
  # Extract residuals from the model
  res <- residuals(log_model)
  
  # Perform ARCH test on the residuals
  arch_test_result <- ArchTest(res, , FALSE)
  
  # Print results
  cat(colnames(log_r_future)[i], "\n")
  print(arch_test_result)
  cat("\n")
}

```

```{r}
# Fit ARIMA model for soya bean meal
arima_soya <- auto.arima(log_r_future[,2])
summary(arima_soya)

# Fit ARIMA model for rice
arima_rice <- auto.arima(log_r_future[,3])
summary(arima_rice)


# Forecast for 10 periods ahead
forecast_soya <- forecast(arima_soya, h = 25)
forecast_rice <- forecast(arima_rice, h = 25)
 
# Plot forecasts
autoplot(forecast_soya, main = "ARIMA Forecast for Soya Bean Log Returns")
autoplot(forecast_rice, main = "ARIMA Forecast for Rice Log Returns")

checkresiduals(arima_soya)
checkresiduals(arima_rice)
```

# GARCH Model evaluation
```{r}
# Function to fit GARCH models with different orders and print AIC/BIC
spec_corn <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(p, q)),
                       mean.model = list(armaOrder = c(0, 0), include.mean = TRUE), distribution.model = "std")
garchfit_corn <- ugarchfit(spec, log_r_future[,4])

spec_wheat <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(p, q)),
                       mean.model = list(armaOrder = c(0, 0), include.mean = TRUE), distribution.model = "std")
garchfit_wheat <- ugarchfit(spec, log_r_future[,5])

print(ugarchforecast(fit = garchfit_corn, n.ahead = 25))
print(ugarchforecast(fit = garchfit_wheat, n.ahead = 25))

plot(ugarchforecast(fit = garchfit_corn, n.ahead = 25), which = 1)
plot(ugarchforecast(fit = garchfit_corn, n.ahead = 25), which = 3)
plot(ugarchforecast(fit = garchfit_wheat, n.ahead = 25), which = 1)
plot(ugarchforecast(fit = garchfit_wheat, n.ahead = 25), which = 3)
```